The Galactic Midplane is Not a Plane: Implications for {\em Gaia}

### 2019-04-22 paper/ms.tex
change to title from meeting discussion
Change range dphi plot to 10% error

%mm substantive Mac Low changes marked so.  Minor changes done silently.

@@

% affiliations
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute,
162 5th Ave., New York, NY 10010, USA}
\newcommand{\penn}{Department of Physics \& Astronomy, University of
Pennsylvania, 209 South 33rd St., Philadelphia, PA 19104, USA}
\newcommand{\amnh}{Department of Astrophysics, American Museum of Natural
History, Central Park West at 79th St., New York, NY 10024, USA}
\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 W
120th St., New York, NY 10027, USA}
\newcommand{\victoria}{Departmeny of Physics \& Astronomy, University of
Victoria, 3800 Finnerty Rd., Victoria, BC, V8P 4HN, Canada}
\newcommand{\nyuphys}{Center for Cosmology and Particle Physics, Department of
Physics, New York University, 726 Broadway, New York, NY 10003, USA}
\newcommand{\nyucds}{Center for Data Science, New York University, 60 Fifth
Ave., New York, NY 10011, USA}
\newcommand{\mpia}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17,
69117 Heidelberg, Germany}

@@

% \correspondingauthor{Angus Beane}
%mm [will this address go away during the review process?.  Do you
%want to give a "current address" as a second footnote?]
\email{abeane@sas.upenn.edu}

@@

Stellar actions, computed for stars from their six-dimensional phase space measurement and
an assumed Galactic potential, are used to label and distinguish orbits. In
principle, actions have the attractive quality of being invariant and thus
good labels of a star's orbit. 
%mm Of course,
    Such analyses are often used to interpret data from the {\em Gaia}
    mission. However, 
inaccurate measurements of the 
phase space position and uncertainties in the assumed Galactic potential
induce an error in the computed action. We show that, in addition to these
complications, a systematic bias in the phase space position induces a
phase-dependence in the actions, which we interpret as 
%mm an 
    a systematic
error. The induced error distribution is
non-Gaussian and bimodal, 
with neither of the modes peaking on
%mm  the true action value. [the error in a Gaussian peaks on zero] 
    the null value.
A midplane error of $\sim30\,\pc$ for a thin disk orbit or
$\sim225\,\pc$ for a thick disk orbit is 
%mm all that is necessary 
      sufficient
to induce a $100\,\%$ error in the vertical action $J_z$.
%mm [fragment?] orbits, respectively. 
Furthermore, we show that the local midplane varies by
$\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
simulations from the FIRE collaboration. Thus, current state-of-the-art action
calculations --- which assume the global and local midplanes are the same ---
include a systematic 
%mm offset in $z$.
   vertical offset variation. 
Variation in the local standard of rest
%mm will induce 
     induces
similar issues. The variation of the midplane must be taken into
account when performing dynamical modelling across
the large regions of the disk
accessible 
%mm with the {\em Gaia} mission. [moved first mention up to top of
%abstract, given that this is in the title.]
    to {\em Gaia}.

@@

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a
revolution 
%mm in the era of {\em Gaia} DR2. 
     as a result of {\em Gaia} Data Release 2 (DR2).  Recent major discoveries include the remnants of a
major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the 2019arXiv190405721A
solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
local substructure infall \citep{2018MNRAS.481.1501B, 2018arXiv180800451L} or
bar buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation
by a dark matter subtructure in the tidal stream GD1
\citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
indicate that the Milky Way's stellar distribution, which demonstrably departs
significantly from axisymmetry, is undergoing phase mixing and dynamical
interactions across a range of spatial and temporal scales.

Underlying the quantitative analysis of many of the mechanisms at work that
give rise to these signatures is the assumption of a global, axisymmetric
Galactocentric coordinate system \citep{2008gady.book.....B} in which the
Sun's position (and therefore the origin of the coordinate system) can be
defined and measured both precisely and accurately. In this system, the center
of the Galaxy is usually taken to be the location of the central supermassive
black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
the Galactic center
%mm is taken to be $\sim 8.1\text{--}8.3\,\kpc$
    can be measured
\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent 
%mm $\sim0.3\%$ measurement from the GRAVITY collaboration
    measurement using near-infrared interferometry of $8.178 \pm
    0.035$~kpc 
\citep{2019arXiv190405721A}. The Sun's vertical distance
from the Galactic midplane, the other independent measurement that sets this
coordinate system, can be determined by identifying where the stellar density
and velocities reach a maximum (effectively the median height of all disk
stars), and is usually taken to be $\sim 25\,\pc$ \citep{2001ApJ...553..184C},
with more recent measurements from {\em Gaia} DR2 placing it at $\sim 20\,\pc$
\citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII
regions in the disk to define the Galactic midplane, leading to slightly
different values ($\sim 5\,\pc$) for the Sun's relative height
\citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
measurements is given by \citet{2016ARAA..54..529B}.

Once a Galactocentric coordinate system has been established and a
six-dimensional (6D) phase
space measurement of a star has been made, it is often desirable to convert
this measurement into action space to concisely summarize its projected orbit,
model the stellar distribution function, or find stars with similar dynamical
properties. Actions are conserved quantities that describe the orbit of a
star under the assumption of regular, bound orbits in a system where the
equations of motion are separable in a particular coordinate system. They are
the 
%mm typical
integral of the Hamiltonian in cylindrical coordinates 
%mm and are given by,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
\eeq
where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
respectively in a cylindrical coordinate system and the $p_i$ are the
conjugate momenta.\footnote{We 
%mm typically [there are no other J_R's in manuscript]
write $J_r$ instead of $J_R$ for
aesthetic reasons.} In a slowly-evolving axisymmetric potential, these actions
are invariant and $J_{\phi} \equiv L_z$
\citep{2008gady.book.....B,2014RvMP...86....1S}.

With the advent of 6D phase-space measurements over a relatively large
($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars'
actions has gained new popularity. One reason is dimensionality reduction---an
individual stellar orbit is concisely described by three actions, as opposed
to six phase space coordinates. Second, under the assumption of a
phase-mixed system, the dynamical properties of a population of stars should
be uniquely a function of their actions and independent of the conjugate
angles, allowing investigation of the relationship between {\em orbital}
properties of stars and other intrinsic, and at least partially invariant,
properties such as age or metallicity. Actions also provide a convenient basis
for constructing models of the stellar distribution function, or for
associating stars with similar dynamical properties (e.g. to determine
membership in moving groups).

@@

The transformation to action space is nonlinear in both the phase-space
position and the potential, and requires relatively strong assumptions about
the system's symmetry. 
%mm For instance, 
    However, 
the Galactic potential is not strictly
axisymmetric; with the vast improvement in the quality of phase-space
measurements due to {\em Gaia}, this assumption appears increasingly
inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this
assumption 
%mm is [counterfactual subjunctive, I believe]
     were
close enough for many purposes, the parameters used in axisymmetric models may
be inadequately constrained by current observations.

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position,
thanks to relatively small-scale variations in the density of the gas and
stellar distribution. Converting positions and velocities of more distant
stars from a heliocentric to a Galactocentric coordinate system introduces a
systematic bias in the $z$ coordinate, and hence the actions computed for
these stars. The further from the solar neighborhood the target star, the more
likely the mismatch will result in large systematic uncertainty, especially in
the vertical action $J_z$. Depending on the type of orbit being studied, this
sets a limit on the ``dynamical solar neighborhood'' over which the local
coordinate frame can be usefully extended for action-space analysis. A similar
argument applies to 
%mm [given that this is down to 0.3% error, seems a bit academic]
   any remaining 
uncertainty in measurements of 
%mm $R_{\odot}$\footnote{Here,
%and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
%Galactic center.}
     the Galactocentric radius of the Sun $R_{\odot}$
and to variations in the local standard of rest (LSR).

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. Lacking an empirical, nonlocal
definition of the global Galactic midplane or a measurement of the expected
midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
we 
%mm attempt to quantify the 
    examine the expected 
azimuthal dependence of the local midplane, and
hence the expected size of the dynamical solar neighborhood, using
cosmological, hydrodynamical, zoom-in simulations of Milky Way-mass galaxies
from the Feedback in Realistic Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2016ApJ...827L..23W,2017arXiv170206148H}.

In Section~\ref{sec:discussion} we discuss the implications of midplane
variations, and the resulting systematic uncertainty in the vertical action,
for action-space analyses beyond the dynamical solar neighborhood. As an
illustration, in Section \ref{sssec:neighborhood} we use these estimates to
consider the feasibility of one science goal that relies on the invariance of
actions---the reconstruction of open clusters across large distances using
``dynamical tagging;'' that is, the association of stars to a birth cluster
based on their similar actions. In light of our findings, we considered the
orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
Gaia} collaboration \citep{2018AA...616A..10G}. In
Table~\ref{tab:real_clusters}, we report the distance to each cluster, the
three actions, and the maximum vertical extent of their orbits
($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we estimate and
discuss the size of the dynamical solar neighborhood for each cluster based on
its mass, to show that action-space tagging is likely still a viable method if
contamination from foreground or background stars is unimportant. We summarize
and conclude in Section~\ref{sec:conclusion}.

@@

@@

\begin{deluxetable*}{ccccccc}
\tablecaption{Description and names of the three orbits considered in this
work.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16
\enddata
\end{deluxetable*}

@@

\begin{figure}
\plotone{fig/schmactions_Jz_hist.pdf}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account.}
\label{fig:Jz_hist}
\end{figure}

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em right} panel
of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
as a vertical dashed line. Here we see that the 
%mm spread 
    errors 
in $J_z$ induced by a
systematic offset in $z$ are non-Gaussian and bimodal. Furthermore, neither of
the modes lie %mm at the true value of $J_z$.
   at null. 
In the case of the thin-disk orbit
({\em lower panel}), we see that, in addition to the prior complications, the
error distribution is not even centered on 
%mm the true value. 
    null. 
This comes about
when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
realizing that in such a case the midplane offset always acts to increase the
value of $J_z$, resulting in a biased estimate of $J_z$.

Empirically describing the error distribution shown in
Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
are not adequate. We therefore elect to measure this error by computing the
95\uth{} percentile minus the 5\uth{} percentile of the distribution of
observed action values. We refer to this quantity as $\Delta J_i$ for each
action value. Because of the bimodality of the error distribution, this
quantity roughly measures the distance between the peaks of the two modes.
Furthermore, this bimodality also implies that $\Delta J_i$ is not very
sensitive to the exact percentiles used. This summary statistic does not
reflect the bias induced when the midplane error is $\gtrsim z_{\text{max}}$.

We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref}
but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report the 95\uth{}
minus 5\uth{} percentile divided by the true action value for the three
different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
we also now consider the effect on the action determined for the thin-disk
(\thincolor) and halo (\halocolor) orbits.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as the 95\uth{}
minus 5\uth{} percentile of the action computed using the offset coordinate
system over the course of the first $\Gyr$, divided by the true action
(computed in the correct coordinate system).

@@

\begin{figure}
\plotone{fig/cluster_offset.pdf}
\caption{The $z$ offset necessary in order for the action error induced by our
midplane effect to be comparable to the intrinsic action dispersion from a
cluster of mass $m_c$. We compute the intrinsic dispersion from
Equation~\ref{eq:action_disp} and convert the dispersion to a $z$ offset using
the result from Figure~\ref{fig:many_orbit_wrong_ref}. We consider the result
for the thin and thick orbit (Table~\ref{tab:orbits}). Nearby open clusters
(Table~\ref{tab:real_clusters}) have orbits closest to the thin orbit .}
\label{fig:cluster_offset}
\end{figure}

>>>>

@@

% affiliations
\newcommand{\cca}{Center for Computational Astrophysics, Flatiron Institute,
162 5th Ave., New York, NY 10010, USA}
\newcommand{\penn}{Department of Physics \& Astronomy, University of
Pennsylvania, 209 South 33rd St., Philadelphia, PA 19104, USA}
\newcommand{\amnh}{Department of Astrophysics, American Museum of Natural
History, Central Park West at 79th St., New York, NY 10024, USA}
\newcommand{\columbia}{Department of Astronomy, Columbia University, 550 W
120th St., New York, NY 10027, USA}
\newcommand{\victoria}{Department of Physics \& Astronomy, University of
Victoria, 3800 Finnerty Rd., Victoria, BC, V8P 4HN, Canada}
\newcommand{\nyuphys}{Center for Cosmology and Particle Physics, Department of
Physics, New York University, 726 Broadway, New York, NY 10003, USA}
\newcommand{\nyucds}{Center for Data Science, New York University, 60 Fifth
Ave., New York, NY 10011, USA}
\newcommand{\mpia}{Max-Planck-Institut f\"ur Astronomie, K\"onigstuhl 17,
69117 Heidelberg, Germany}

@@

% \correspondingauthor{Angus Beane}
\email{abeane@sas.upenn.edu}

@@

Stellar actions, computed for stars from their six-dimensional phase space measurement and
an assumed Galactic potential, are used to label and distinguish orbits. In
principle, actions have the attractive quality of being invariant and thus
good labels of a star's orbit. Of course, inaccurate measurements of the 
phase space position and uncertainities in the assumed Galactic potential
induce an error in the computed action. We show that, in addition to these
complications, a systematic bias in the phase space position induces a
phase-dependence in the actions, which we interpret as an error. The induced error distribution is
non-Gaussian and bimodal, 
with neither of the modes peaking on the true action value. 
A midplane error of $\sim6\,\pc$ for a thin disk orbit or
$\sim50\,\pc$ for a thick disk orbit is all that is necessary to
induce a $10\,\%$ error in the vertical action $J_z$.
Furthermore, we show that the local midplane varies by
$\sim100\,\pc$ at the Solar circle in Milky Way-mass cosmological zoom-in
simulations from the FIRE collaboration. Thus, current state-of-the-art action
calculations --- which assume the global and local midplanes are the same ---
include a systematic offset in $z$. Variation in the local standard of rest
will induce similar issues. The variation of the midplane must be taken into
account when performing dynamical modelling across large regions of the disk
accessible with the {\em Gaia} mission.

@@

\section{Introduction} \label{sec:intro}
Our understanding of the Milky Way is currently undergoing a revolution in the
era of {\em Gaia} DR2. Recent major discoveries include the remnants of a
major merger \citep{2018ApJ...860L..11K, 2018Natur.563...85H,
2018arXiv180704290L, 2019MNRAS.482.3426M}, a phase-space ``spiral'' in the
solar neighborhood \citep{2018Natur.561..360A} possibly indicating either
local substructure infall \citep{2018MNRAS.481.1501B, 2018arXiv180800451L} or
bar buckling \citep{2018arXiv181109205K}, and a gap suggestive of perturbation
by a dark matter subtructure in the tidal stream GD1
\citep{2018ApJ...863L..20P, 2018arXiv181103631B}. These discoveries all
indicate that the Milky Way's stellar distribution, which demonstrably departs
significantly from axisymmetry, is undergoing phase mixing and dynamical
interactions across a range of spatial and temporal scales.

Underlying the quantitative analysis of many of the mechanisms at work that
give rise to these signatures is the assumption of a global, axisymmetric
Galactocentric coordinate system \citep{2008gady.book.....B} in which the
Sun's position (and therefore the origin of the coordinate system) can be
defined and measured both precisely and accurately. In this system, the center
of the Galaxy is usually taken to be the location of the central supermassive
black hole, Sgr~A\textsuperscript{*} \citep[e.g.][]{2004ApJ...616..872R}. From
stellar motions near Sgr~A\textsuperscript{*}, the distance from the Sun to
the Galactic center is taken to be $\sim 8.1\text{--}8.3\,\kpc$
\citep{2009ApJ...692.1075G,2018AA...615L..15G}, with a recent $\sim0.3\%$ measurement from the GRAVITY collaboration \citep{2019arXiv190405721A}. The Sun's vertical distance
from the Galactic midplane, the other independent measurement that sets this
coordinate system, can be determined by identifying where the stellar density
and velocities reach a maximum (effectively the median height of all disk
stars), and is usually taken to be $\sim 25\,\pc$ \citep{2001ApJ...553..184C},
with more recent measurements from {\em Gaia} DR2 placing it at $\sim 20\,\pc$
\citep{2019MNRAS.482.1417B}. Another strategy is to use the cold gas or HII
regions in the disk to define the Galactic midplane, leading to slightly
different values ($\sim 5\,\pc$) for the Sun's relative height
\citep[e.g.][]{2019ApJ...871..145A}. A pre-{\em Gaia} review of these
measurements is given by \citet{2016ARAA..54..529B}.

Once a Galactocentric coordinate system has been established and a 6D phase
space measurement of a star has been made, it is often desirable to convert
this measurement into action space to concisely summarize its projected orbit,
model the stellar distribution function, or find stars with similar dynamical
properties. Actions are conserved quantities which describe the orbit of a
star under the assumption of regular, bound orbits in a system where the
equations of motion are separable in a particular coordinate system. They are
the typical integral of the Hamiltonian in cylindrical coordinates and are
given by,
\beq\label{eq:actions}
J_i \equiv
\frac{1}{2\pi} \oint_{\text{orbit}}p_i\,\text{d}x_i\text{,}
\eeq
where $i=R,z,\phi$ are the radial, vertical, and azimuthal coordinates
respectively in a cylindrical coordinate system and the $p_i$ are the
conjugate momenta.\footnote{We typically write $J_r$ instead of $J_R$ for
aesthetic reasons.} In a slowly-evolving axisymmetric potential, these actions
are invariant and $J_{\phi} \equiv L_z$
\citep{2008gady.book.....B,2014RvMP...86....1S}.

With the advent of 6D phase-space measurements over a relatively large
($\gtrsim 2$ kpc) volume from the {\em Gaia} satellite, the study of stars'
actions has gained new popularity. One reason is dimensionality reduction---an
individual stellar orbit is concisely described by three actions, as opposed
to six dimensions of phase space. Second, under the assumption of a
phase-mixed system, the dynamical properties of a population of stars should
be uniquely a function of their actions and independent of the conjugate
angles, allowing investigation of the relationship between {\em orbital}
properties of stars and other intrinsic, and at least partially invariant,
properties such as age or metallicity. Actions also provide a convenient basis
for constructing models of the stellar distribution function, or for
associating stars with similar dynamical properties (e.g. to determine
membership in moving groups).

@@

The transformation to action space is nonlinear in both the phase-space
position and the potential, and requires relatively strong assumptions about
the system's symmetry. For instance, the Galactic potential is not strictly
axisymmetric; with the vast improvement in the quality of phase-space
measurements due to {\em Gaia}, this assumption appears increasingly
inadequate \citep[e.g.][]{2018Natur.561..360A}. Even if this assumption is
close enough for many purposes, the parameters used in axisymmetric models may
be inadequately constrained by current observations.

In this work we discuss how the imposition of a global axisymmetric coordinate
system based on a local measurement of the midplane location and
Galactocentric distance introduces systematic error in the computation of
actions, especially for stars in the thin disk. Our main point is that the
``local'' Galactic midplane is not the same as the ``global'' Galactic
midplane. For stars relatively far from the Sun, the local midplane location
will not agree with our local midplane extrapolated onto their position,
thanks to relatively small-scale variations in the density of the gas and
stellar distribution. Converting positions and velocities of more distant
stars from a heliocentric to a Galactocentric coordinate system introduces a
systematic bias in the $z$ coordinate, and hence the actions computed for
these stars. The further from the solar neighborhood the target star, the more
likely the mismatch will result in large systematic uncertainty, especially in
the vertical action $J_z$. Depending on the type of orbit being studied, this
sets a limit on the ``dynamical solar neighborhood'' over which the local
coordinate frame can be usefully extended for action-space analysis. A similar
argument applies to uncertainty in measurements of $R_{\odot}$\footnote{Here,
and throughout, we use $R_{\odot}$ to mean the distance from the Sun to the
Galactic center.} and variations in the local standard of rest (LSR).

In Section~\ref{sec:ref_frame}, we describe the general impact coordinate
system errors have on the measured actions. Lacking an empirical, nonlocal
definition of the global Galactic midplane or a measurement of the expected
midplane variation in the Milky Way itself,  in Section~\ref{sec:local_fire}
we attempt to quantify the azimuthal dependence of the local midplane, and
hence the expected size of the dynamical solar neighborhood, using
cosmological-hydrodynamical zoom-in simulations of Milky Way-mass galaxies
from the Feedback in Realistic Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}}
\citep{2016ApJ...827L..23W,2017arXiv170206148H}.

In Section~\ref{sec:discussion} we discuss the implications of midplane
variations, and the resulting systematic uncertainty in the vertical action,
for action-space analyses beyond the dynamical solar neighborhood. As an
illustration, in Section \ref{sssec:neighborhood} we use these estimates to
consider the feasibility of one science goal that relies on the invariance of
actions --- the reconstruction of open clusters across large distances using
``dynamical tagging;'' that is, the association of stars to a birth cluster
based on their similar actions. In light of our findings, we considered the
orbits of the nine open clusters within $250\,\pc$ as reported by the {\em
Gaia} collaboration \citep{2018AA...616A..10G}. In
Table~\ref{tab:real_clusters}, we report the distance to each cluster, the
three actions, and the maximum vertical extent of their orbits
($z_{\text{max}}$). In Section~\ref{sssec:reconstruction} we estimate and
discuss the size of the dynamical solar neighborhood for each cluster based on
its mass, to show that action-space tagging is likely still a viable method if
contamination from foreground or background stars is unimportant. We summarize
and conclude in Section~\ref{sec:conclusion}.

@@

\subsection{Epicyclic Approximation} \label{ssec:epi_action}
Before turning to numerical methods, we first write down analytically the
induced action error from either a midplane offset or a Galactocentric radius
offset. We make the epicyclic approximation, which assumes that the motion in
the $z$ and $R$ components of the orbit are decoupled and follow a simple
harmonic oscillator. This approximation should be an excellent description of
thin disk-like orbits and a good description of thick disk-like orbits. We
also make the assumption of a perfectly flat circular velocity curve with
velocity $v_c$.

Under this approximation, we can write down the vertical and radial components
of the orbits as,
\beq\label{eq:orbits_epi}
\begin{split}
z(t) &= A_z \sin{(\kappa t + \alpha)}\\
R(t) &= A_R \sin{(\nu t + \beta)}\text{,}
\end{split}
\eeq
where $\kappa$ and $\nu$ are the vertical and radial (or epicyclic)
frequencies, $A_z$ and $A_R$ are the amplitudes of the stellar motion in the
vertical and radial coordinates, and $\alpha$ and $\beta$ are arbitrary
constants. We can write down the actions in this approximation as
\citep[Section~3.5.3b][]{2008gady.book.....B},
\beq\label{eq:actions_epi}
\begin{split}
L_z &= R_g v_c \\
J_r &= \kappa A_R^2 \\
J_z &= \nu A_z^2\text{,}
\end{split}
\eeq
where $R_g$ is the guiding radius of the orbit.

An induced error in $z$ of $\Delta z$ will cause $\Delta A_z \sim \Delta z$.
Similarly, an induced error in $R$ will cause $\Delta A_R \sim \Delta R$. If
we assume that $\kappa$ and $\nu$ are constant (\Gus{justify?}), then we have
that
\beq\label{eq:actions_epi_error}
\begin{split}
\Delta L_z = \Delta R v_c &\implies \frac{\Delta L_z}{L_z} = \frac{\Delta R}{R} \\
\Delta J_r = 2 \kappa \Delta A_R &\implies \frac{\Delta J_r}{J_r} = \frac{2\Delta R}{A_R} \\
\Delta J_z = 2 \nu \Delta A_z &\implies \frac{\Delta J_z}{J_z} = \frac{2\Delta z}{A_z}\text{.}
\end{split}
\eeq
Therefore, errors in each action are linearly proportional to the errors in
either the midplane or cylindrical radius.

There is a slight complication in the formula for computing $\Delta J_r/J_r$ and $\Delta L_z/L_z$.
In a moment we will consider the effect of an inaccuracy in the distance from
the Sun to the Galactic center on actions. However, this inaccuracy induces an
error in $x$, or $\Delta x$. This is not the same as an inaccuracy in $\Delta
R$, but they are related by the following argument. An error $\Delta x$ will
induce an erroneous radius $R_{\text{err}}$ related by the formula,
\beq
(x+\Delta x)^2 + y^2 = R_{\text{err}}^2\text{.}
\eeq
Keeping only terms to first order in $\Delta x$, we have that,
\beq
\begin{split}
R_{\text{err}}^2 &= R^2 - 2 R \cos{\theta} \Delta x \\
\implies \Delta R &\equiv \abs{R_{\text{err}} - R} = \abs{\cos{\theta}} \Delta x\text{.}
\end{split}
\eeq
Averaging over the circle, we therefore have that
\beq
\avg{\Delta R} = \frac{2}{\pi} \Delta x\text{.}
\eeq
Therefore, the error in $J_r$ and $L_z$ are related to $\Delta x$ by,
\beq\label{eq:Jr_epi_dx}
\begin{split}
\frac{\Delta J_r}{J_r} &= \frac{4 \Delta x}{\pi A_R} \\
\frac{\Delta L_z}{L_z} &= \frac{2 \Delta x}{\pi R}\text{.}
\end{split}
\eeq

We will check these formulae against numerical tests next.

@@

\begin{deluxetable*}{cccccccc}
\tablecaption{Description and names of the three orbits considered in this
work. For the last two columns: $z_{\text{max}}$ the maximum height of the
orbit while $A_R$ is the magnitude of the radial excursions of the orbit,
computed as one half the maximum radius minus the minimum
radius.\label{tab:orbits}}
\tablehead{\colhead{name} & \colhead{initial position} & \colhead{initial
velocity} & \colhead{$J_r$} &
\colhead{$L_z$} & \colhead{$J_z$} & \colhead{$z_{\text{max}}$} &
\colhead{$A_R$}\\ \colhead{ } &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{km}/\mathrm{s}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc\,km\,s^{-1}}$} & \colhead{$\mathrm{kpc\,km\,s^{-1}}$} &
\colhead{$\mathrm{kpc}$} & \colhead{$\mathrm{kpc}$}}
\startdata 
thin-disk & $(8, 0, 0)$ & $(0, -190, 10)$ & 40.3 & -1520 & 0.69 & 0.12 & 1.29 \\
thick-disk & $(8, 0, 0)$ & $(0, -190, 50)$ & 32.5 & -1520 & 23.0 & 0.85 & 1.19 \\ 
halo & $(8, 0, 0)$ & $(0, -190, 190)$ & 32.8 & -1520 & 529.1 & 6.16 & 2.34
\enddata
\end{deluxetable*}

@@

\begin{figure}
\plotone{fig/schmactions_Jz_hist.pdf}
\caption{A histogram of the observed values in $J_z$ for the thick-disk orbit
assuming a $z$ offset of $100\,\pc$. One can see that if the observed $z$
values have a bias (from e.g. an incorrectly computed midplane), then the
induced error distribution in $J_z$ is decidedly non-Gaussian. Therefore, any
sort of error propagation must take this into account. We also plot one half
the 95\uth{} percentile minus the 5\uth{} percentile of each distribution as a
horizontal arrow anchored on the true $J_z$ value. We call this $\Delta J_z$
and will use it (along with the similarly defined $\Delta J_r$ and $\Delta
L_z$) to empirically describe the error distribution. We see that $\Delta J_z$
roughly corresponds to the distance from the true $J_z$ value to one of the
modes of the error distribution.}
\label{fig:Jz_hist}
\end{figure}

In Figure~\ref{fig:Jz_hist} we plot a histogram of the observed values of
$J_z$ for the thick-disk orbit ({\em upper panel}) and thin-disk orbit ({\em
lower panel}) assuming a $z$ offset of $100\,\pc$ (i.e. the {\em right} panel
of Figure~\ref{fig:one_orbit_wrong_ref}). The true value of $J_z$ is plotted
as a vertical dashed line. Here we see that the spread in $J_z$ induced by a
systematic offset in $z$ is non-Gaussian and bimodal. Furthermore, neither of
the modes lie at the true value of $J_z$. In the case of the thin-disk orbit
({\em lower panel}), we see that, in addition to the prior complications, the
error distribution is not even centered on the true value. This comes about
when the midplane error is $\gtrsim z_{\text{max}}$, and can be understood by
realizing that in such a case the midplane offset always acts to increase the
value of $J_z$, resulting in a biased estimate of $J_z$.

Empirically describing the error distribution shown in
Figure~\ref{fig:Jz_hist} is difficult. Of course, Gaussian summary statistics
are not adequate. We therefore elect to measure this error by computing one half the
95\uth{} percentile minus the 5\uth{} percentile of the distribution of
observed action values. We refer to this quantity as $\Delta J_i$ for each
action value. We plot this quantity in Figure~\ref{fig:Jz_hist} as a horizontal arrow anchored on the true action value. Because of the bimodality of the error distribution, this
quantity roughly measures the distance from the true action value to the peak of one of the modes.
Furthermore, this bimodality also implies that $\Delta J_i$ is not very
sensitive to the exact percentiles used. This summary statistic does not
reflect the bias induced when the midplane error is $\gtrsim z_{\text{max}}$.

We now repeat the same procedure as in Figure~\ref{fig:one_orbit_wrong_ref}
but for systematic offsets between $0$ and $500\,\pc$ in the $z$ and $x$
components. In Figure~\ref{fig:many_orbit_wrong_ref}, we report one half the 95\uth{}
minus 5\uth{} percentile divided by the true action value ($\Delta J_i/J_i$) for the three
different fiducial orbits in Table \ref{tab:orbits}. The thick-disk
(\thickcolor) orbit is the one from Figure~\ref{fig:one_orbit_wrong_ref}, but
we also now consider the effect on the action determined for the thin-disk
(\thincolor) and halo (\halocolor) orbits.

The top row of Figure~\ref{fig:many_orbit_wrong_ref} shows the spread induced
in each action for an offset in the $z$-component (i.e. the Galactic
midplane). In the {\em bottom row} we consider offsets in the $x$ component
(i.e. the Solar radius). The {\em left}, {\em middle}, and {\em right} columns
show the fractional spread in the values computed for $J_z$, $L_z$, and $J_r$,
respectively. We compute the fractional spread $\Delta J_i/J_i$ as one half the 95\uth{}
minus 5\uth{} percentile of the action computed using the offset coordinate
system over the course of the first $\Gyr$, divided by the true action
(computed in the correct coordinate system).

@@

\begin{figure}
\plotone{fig/cluster_offset.pdf}
\caption{The $z$ offset necessary in order for the action error induced by our
midplane effect to be comparable to the intrinsic action dispersion from a
cluster of mass $m_c$. We compute the intrinsic dispersion from
Equation~\ref{eq:action_disp} and convert the dispersion to a $z$ offset using
the result from Figure~\ref{fig:many_orbit_wrong_ref}. We consider the result
for the thin and thick orbit (Table~\ref{tab:orbits}). Nearby open clusters
(Table~\ref{tab:real_clusters}) have orbits closest to the thin orbit.}
\label{fig:cluster_offset}
\end{figure}
